\documentclass[aspectratio=169,11pt]{beamer}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{ragged2e}
\usepackage{colortbl}
\usepackage{array}
\usepackage{hyperref}
\usepackage{adjustbox}

% --- Color Palette: Warm Professional ---
\definecolor{DeepNavy}{HTML}{2E4057}
\definecolor{Teal}{HTML}{048A81}
\definecolor{WarmOrange}{HTML}{E85D04}
\definecolor{SoftPurple}{HTML}{9D4EDD}
\definecolor{WarmGray}{HTML}{6C757D}
\definecolor{LightGray}{HTML}{E9ECEF}
\definecolor{Cream}{HTML}{FBF8F1}
\definecolor{DeepRed}{HTML}{D62828}
\definecolor{Gold}{HTML}{D4A03A}
\definecolor{SoftWhite}{HTML}{FAFAFA}

% Beamer colors
\setbeamercolor{normal text}{fg=DeepNavy, bg=SoftWhite}
\setbeamercolor{structure}{fg=DeepNavy}
\setbeamercolor{alerted text}{fg=DeepRed}
\setbeamercolor{example text}{fg=Teal}
\setbeamercolor{frametitle}{fg=DeepNavy, bg=SoftWhite}
\setbeamercolor{title}{fg=DeepNavy}
\setbeamercolor{subtitle}{fg=WarmGray}
\setbeamercolor{block title}{bg=DeepNavy, fg=SoftWhite}
\setbeamercolor{block body}{bg=LightGray, fg=DeepNavy}
\setbeamercolor{itemize item}{fg=WarmOrange}
\setbeamercolor{itemize subitem}{fg=Gold}

% Fonts
\usefonttheme{professionalfonts}
\setbeamerfont{title}{size=\huge, series=\bfseries}
\setbeamerfont{frametitle}{size=\Large, series=\bfseries}

% Strip chrome
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{headline}{}

% Minimal footline
\setbeamertemplate{footline}{%
    \hfill
    \begin{beamercolorbox}[wd=3cm,ht=2.5ex,dp=1ex,right,rightskip=0.6cm]{page number}
        \usebeamercolor[fg]{WarmGray}\scriptsize\insertframenumber
    \end{beamercolorbox}
    \vspace{0.3cm}
}

% Bullets
\setbeamertemplate{itemize item}{\footnotesize\raisebox{0.3ex}{\tikz\fill[WarmOrange] (0,0) circle (0.45ex);}}
\setbeamertemplate{itemize subitem}{\footnotesize\raisebox{0.3ex}{\tikz\fill[Gold] (0,0) circle (0.35ex);}}

% Suppress beamer's cosmetic underfull \hbox warnings on flushright/RaggedRight lines.
% Overfull warnings remain enforced.
\hbadness=10000

% Custom commands
\newcommand{\transitionslide}[2]{%
    {\setbeamercolor{normal text}{fg=SoftWhite,bg=DeepNavy}
    \begin{frame}[plain]\vfill\begin{center}
        {\huge\bfseries\textcolor{Gold}{#1}}\\[0.6cm]
        {\large\textcolor{LightGray}{#2}}
    \end{center}\vfill\end{frame}}
}

\title{\textbf{The Direct Answer Dilemma}}
\subtitle{How AI Search Closes the Attribution Loop \textemdash{} and Who Pays for It}
\author{Audit synthesis}
\date{}

\begin{document}

% ---------- TITLE ----------
{\setbeamertemplate{footline}{}
\begin{frame}[plain]
\titlepage
\vspace{-0.6cm}
\begin{center}
\footnotesize\textcolor{WarmGray}{%
Potthast et al.\ (2020), \emph{SIGIR Forum} \quad\textbar\quad
Liu, Zhang \& Liang (2023), \emph{arXiv:2304.09848}\\
Jazwinska \& Chandrasekar (2024, 2025), \emph{Tow Center / CJR} \quad\textbar\quad
Aggarwal et al.\ (2024), \emph{KDD '24}\\
Khosravi et al.\ (2026), \emph{arXiv preprint} \quad\textbar\quad
Larsen et al.\ (2025) \quad\textbar\quad
Liang et al.\ (2025), \emph{Patterns}}
\end{center}
\end{frame}}

% ---------- OPENING PROVOCATION ----------
\begin{frame}[plain]
\vfill
\begin{center}
{\Huge\bfseries\textcolor{DeepRed}{60\%}}\\[0.4cm]
{\large\textcolor{DeepNavy}{of citation queries answered \emph{wrong} by eight AI search engines.}}\\[0.2cm]
{\large\textcolor{DeepNavy}{Grok 3: 94\% wrong. Most wrong answers came with no hedging.}}\\[1.0cm]
{\Large\textcolor{WarmGray}{The interface earned the user's trust. The citations did not.}}
\end{center}
\vfill
\begin{flushright}
\scriptsize\textcolor{WarmGray}{Jazwinska \& Chandrasekar (2025), \emph{Tow Center / CJR}; 1{,}600 queries across 8 chatbots.}
\end{flushright}
\end{frame}

% ---------- FRAMEWORK ----------
\begin{frame}{Potthast (2020) named four risks of the direct-answer interface \textemdash{} all four now have data}
\begin{center}
\renewcommand{\arraystretch}{1.05}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{4.4cm}>{\RaggedRight\arraybackslash}p{4.0cm}>{\RaggedRight\arraybackslash}p{3.8cm}@{}}
\toprule
\textbf{Potthast risk (2020)} & \textbf{Mechanism} & \textbf{Empirical confirmation} \\
\midrule
\textbf{Misattribution} & Answer drops the source & \alert{Liu 23; Jazwinska 24/25} \\
\textbf{Source erosion} & Click never happens & \alert{Khosravi 2026} \\
\textbf{Market distortion} & Ad auction reshapes & \alert{Larsen 2025} \\
\textbf{Loop pollution} & Synthetic text re-enters web & \alert{Aggarwal 24; Liang 25} \\
\bottomrule
\end{tabular}
\end{center}
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Potthast, Hagen \& Stein (2020), \emph{SIGIR Forum} 54(1):14.}
\end{flushright}
\end{frame}

% ---------- TRANSITION 1 ----------
\transitionslide{Citation is broken at the source}{Direct-answer interfaces grade their own homework.}

% ============================================================
%  SECTION A \textemdash{} CITATION INTEGRITY (LIU, JAZWINSKA \times 2)
% ============================================================

% ---------- LIU 2023 ----------
\begin{frame}{The better the answer reads, the worse it cites \textemdash{} r = $-0.96$}
\vspace{0.1cm}
\begin{center}
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{4.6cm}p{3.4cm}p{3.4cm}@{}}
\toprule
\textbf{Engine} & \textbf{Citation recall} & \textbf{Citation precision} \\
\midrule
Bing Chat & 58.7\% & 89.5\% \\
NeevaAI & 67.6\% & 72.0\% \\
perplexity.ai & 68.7\% & 72.7\% \\
YouChat & 11.1\% & 63.6\% \\
\midrule
\textbf{Mean (4 engines)} & \alert{\textbf{51.5\%}} & \alert{\textbf{74.5\%}} \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.2cm}
\textbf{Across 1{,}450 statements:} Pearson \(r = -0.96\) between citation precision and the perceived fluency/utility of the answer. Better-sounding answers cite worse.\\[0.2cm]
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Liu, Zhang \& Liang (2023), Tables~3\textendash 4; human eval over four generative search engines, May 2023.}
\end{flushright}
\end{frame}

% ---------- JAZWINSKA 2024 ----------
\begin{frame}{ChatGPT Search misattributed 153 of 200 publisher quotes \textemdash{} and declined only 7 times}
\vspace{0.1cm}
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{5.0cm}p{2.8cm}p{3.4cm}@{}}
\toprule
\textbf{Response category} & \textbf{Count / 200} & \textbf{Share} \\
\midrule
Correct (publisher + date + URL) & 47 & 23.5\% \\
Partially correct & 57 & 28.5\% \\
\textbf{Wrong} & \alert{89} & \alert{44.5\%} \\
Declined / unsure & 7 & 3.5\% \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.2cm}
Even publishers \emph{with} OpenAI licensing deals were misattributed. NYT (in active lawsuit, blocked) was cited via plagiarized copies.\\[0.2cm]
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Jazwinska \& Chandrasekar (2024), \emph{Tow Center / CJR}, Nov.\ 2024; 200 quotes, 20 publishers.}
\end{flushright}
\end{frame}

% ---------- JAZWINSKA 2025 ----------
\begin{frame}{Across 8 chatbots, premium tiers are \emph{more} confidently wrong}
\vspace{0.1cm}
\begin{center}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{4.6cm}p{2.8cm}p{4.0cm}@{}}
\toprule
\textbf{Engine} & \textbf{Inaccuracy} & \textbf{Behavior} \\
\midrule
Perplexity (free) & \textcolor{Teal}{\textbf{37\%}} & best of eight \\
ChatGPT Search & 67\% & hedged 15/200 \\
Perplexity Pro (paid) & higher & decline rate falls \\
Gemini & high & fabricated URLs frequent \\
DeepSeek & high & misattributed 115/200 \\
\textbf{Grok 3 (paid)} & \alert{\textbf{94\%}} & 154/200 broken URLs \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.1cm}
Paid tiers answer \emph{more} prompts confidently \textemdash{} so they make \emph{more} confident errors. Robots.txt routinely violated.\\
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Jazwinska \& Chandrasekar (2025), \emph{CJR}, Mar.\ 2025; 1{,}600 queries (20 publishers $\times$ 10 quotes $\times$ 8 chatbots).}
\end{flushright}
\end{frame}

% ---------- TRANSITION 2 ----------
\transitionslide{The economic price falls on the source}{Direct answers are paid for in someone else's traffic.}

% ============================================================
%  SECTION B \textemdash{} ECONOMIC SPILLOVER (KHOSRAVI, LARSEN)
% ============================================================

% ---------- KHOSRAVI 2026 ----------
\begin{frame}{AI summaries cut Wikipedia pageviews $\sim$15\%, costing \$35.7M\textendash \$102.1M per year}
\vspace{0.1cm}
\begin{center}
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{5.0cm}p{3.6cm}p{2.6cm}@{}}
\toprule
\textbf{Outcome} & \textbf{Estimate} & \textbf{Source} \\
\midrule
Pageview decline (post-AI summaries) & \alert{$-15\%$} & DiD, 2024\textendash 25 \\
Annual donation revenue at risk & \$35.7M\textendash \$102.1M & extrapolation \\
Editor-recruitment funnel & weaker & traffic-volunteer link \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.2cm}
\textbf{The first feedback loop:} fewer readers $\rightarrow$ fewer donations and editors $\rightarrow$ a thinner Wikipedia $\rightarrow$ a thinner training corpus.\\
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Khosravi et al.\ (2026); pageview DiD around AI Overviews / Search GPT rollouts.}
\end{flushright}
\end{frame}

% ---------- LARSEN 2025 ----------
\begin{frame}{Sponsored-search auctions splinter: short-query CPC $+1.5\%$, long-query CPC $-2.5\%$}
\vspace{0.1cm}
\begin{center}
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{5.0cm}p{3.0cm}p{3.0cm}@{}}
\toprule
\textbf{Query length} & \textbf{$\Delta$ CPC (BERT)} & \textbf{Mechanism} \\
\midrule
Short (1\textendash 2 tokens) & \alert{$+1.5\%$} & better intent match $\rightarrow$ tighter rivalry \\
Long-tail (5+ tokens) & \alert{$-2.5\%$} & query absorbed by AI answer \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.2cm}
\textbf{The second feedback loop:} the highest-intent traffic that publishers monetized via long-tail SEO is exactly what AI answers consume \textemdash{} so its auction price falls.\\
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Larsen et al.\ (2025); 11{,}949 queries $\times$ 24 months SEMRush panel; BERT rollout DiD.}
\end{flushright}
\end{frame}

% ---------- TRANSITION 3 ----------
\transitionslide{Creators adapt \textemdash{} and the well gets polluted}{Optimization moves on. The optimization target is now an AI.}

% ============================================================
%  SECTION C \textemdash{} ADAPTATION AND LOOP POLLUTION (AGGARWAL, LIANG)
% ============================================================

% ---------- AGGARWAL 2024 ----------
\begin{frame}{Adding a single quotation lifts AI-engine visibility by 41\%}
\vspace{0.1cm}
\begin{center}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{4.8cm}p{2.6cm}p{3.6cm}@{}}
\toprule
\textbf{GEO tactic} & \textbf{$\Delta$ visibility} & \textbf{Cost to publisher} \\
\midrule
\textbf{Quotation Addition} & \alert{$+41\%$} & low (one edit) \\
Statistics Addition & $+32\%$ & low \\
Authoritative tone & $+30\%$ & stylistic only \\
Citing Sources & $+27\%$ & low \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.2cm}
\textbf{Generative Engine Optimization} (GEO) is now a discipline. The new ranking signal rewards \emph{quotability}, not authority \textemdash{} so creators feed the engine what it likes.\\
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Aggarwal et al.\ (2024), \emph{KDD '24}; benchmark of 9 GEO tactics on 10K queries.}
\end{flushright}
\end{frame}

% ---------- LIANG 2025 ----------
\begin{frame}{By late 2024, $\sim$24\% of corporate press releases and $\sim$18\% of CFPB complaints are LLM-written}
\vspace{0.1cm}
\begin{center}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{4.6cm}p{2.6cm}p{3.6cm}@{}}
\toprule
\textbf{Corpus} & \textbf{LLM share} & \textbf{Sample (N)} \\
\midrule
Corporate press releases & \alert{$\sim$24\%} & 537{,}413 \\
Consumer complaints (CFPB) & $\sim$18\% & 687{,}241 \\
UN press releases & $\sim$14\% & 15{,}919 \\
Small-firm job postings & $\sim$10\textendash 15\% & 1.44M \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.1cm}
Distributional MLE; prediction error \(<3.3\%\); \emph{lower bound} (heavy edits invisible). The corpus on which the next models train is increasingly authored \emph{by} models.
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Liang et al.\ (2025), \emph{Patterns}; four corpora, Jan 2019\textendash Sep 2024.}
\end{flushright}
\end{frame}

% ---------- DEVIL'S ADVOCATE ----------
\begin{frame}{Devil's advocate: aren't most queries the kind where citations don't matter?}
\begin{block}{The strongest objection}
For ``what time is the show,'' ``capital of Peru,'' ``recipe for risotto,'' AI answers users more cheaply than ten blue links \textemdash{} and a wrong citation costs nothing.
\end{block}
\textbf{Reply.} Three problems with the carve-out:
\begin{itemize}
\item Liu (2023): \(r = -0.96\) between fluency and citation precision \textemdash{} users have \emph{no signal} that flips them from ``trust'' to ``verify.''
\item Jazwinska (2025): premium-tier engines decline less, so confident wrongness rises with willingness to pay.
\item Khosravi (2026) + Larsen (2025): even ``trivial'' queries strip the long-tail traffic that funds the verifiable corpus.
\end{itemize}
The cheap query subsidises the expensive one. Sever the cross-subsidy, both fail.
\end{frame}

% ---------- SYNTHESIS ----------
\begin{frame}{Potthast's prophecy is now a self-reinforcing loop with five turns of the crank}
\begin{center}
\renewcommand{\arraystretch}{1.0}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{0.8cm}>{\RaggedRight\arraybackslash}p{3.8cm}>{\RaggedRight\arraybackslash}p{7.0cm}@{}}
\toprule
\textbf{\#} & \textbf{Turn} & \textbf{Empirical anchor} \\
\midrule
1 & Citations decay & Liu 2023: precision\,74.5\%, recall\,51.5\%, $r\!=\!-0.96$ \\
2 & Misattribution scales & Jazwinska 24/25: 60\%\textendash 94\% wrong (8 engines) \\
3 & Page views collapse & Khosravi 2026: $-15\%$ Wikipedia traffic \\
4 & Ad auctions reprice & Larsen 2025: $-2.5\%$ long-tail CPC \\
5 & Synthetic content fills gap & Aggarwal 24 (GEO $+41\%$); Liang 25 ($\sim$24\% PR) \\
\bottomrule
\end{tabular}
\end{center}
The training corpus that produced the model is being eroded by the model. Each turn cheapens the next.
\end{frame}

% ---------- CLOSING ----------
{\setbeamercolor{normal text}{fg=SoftWhite,bg=DeepNavy}
\setbeamertemplate{footline}{}
\begin{frame}[plain]
\vfill
\begin{center}
{\Huge\bfseries\textcolor{Gold}{Direct answers are not free.}}\\[0.6cm]
{\Large\textcolor{LightGray}{They are an externality on the web that produced them.}}
\end{center}
\vfill
\end{frame}}

\end{document}
