\documentclass[aspectratio=169,11pt]{beamer}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{ragged2e}
\usepackage{colortbl}
\usepackage{array}
\usepackage{hyperref}
\usepackage{adjustbox}

% --- Color Palette: Warm Professional ---
\definecolor{DeepNavy}{HTML}{2E4057}
\definecolor{Teal}{HTML}{048A81}
\definecolor{WarmOrange}{HTML}{E85D04}
\definecolor{SoftPurple}{HTML}{9D4EDD}
\definecolor{WarmGray}{HTML}{6C757D}
\definecolor{LightGray}{HTML}{E9ECEF}
\definecolor{Cream}{HTML}{FBF8F1}
\definecolor{DeepRed}{HTML}{D62828}
\definecolor{Gold}{HTML}{D4A03A}
\definecolor{SoftWhite}{HTML}{FAFAFA}

% Beamer colors
\setbeamercolor{normal text}{fg=DeepNavy, bg=SoftWhite}
\setbeamercolor{structure}{fg=DeepNavy}
\setbeamercolor{alerted text}{fg=DeepRed}
\setbeamercolor{example text}{fg=Teal}
\setbeamercolor{frametitle}{fg=DeepNavy, bg=SoftWhite}
\setbeamercolor{title}{fg=DeepNavy}
\setbeamercolor{subtitle}{fg=WarmGray}
\setbeamercolor{block title}{bg=DeepNavy, fg=SoftWhite}
\setbeamercolor{block body}{bg=LightGray, fg=DeepNavy}
\setbeamercolor{itemize item}{fg=WarmOrange}
\setbeamercolor{itemize subitem}{fg=Gold}

% Fonts
\usefonttheme{professionalfonts}
\setbeamerfont{title}{size=\huge, series=\bfseries}
\setbeamerfont{frametitle}{size=\Large, series=\bfseries}

% Strip chrome
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{headline}{}

% Minimal footline
\setbeamertemplate{footline}{%
    \hfill
    \begin{beamercolorbox}[wd=3cm,ht=2.5ex,dp=1ex,right,rightskip=0.6cm]{page number}
        \usebeamercolor[fg]{WarmGray}\scriptsize\insertframenumber
    \end{beamercolorbox}
    \vspace{0.3cm}
}

% Bullets
\setbeamertemplate{itemize item}{\footnotesize\raisebox{0.3ex}{\tikz\fill[WarmOrange] (0,0) circle (0.45ex);}}
\setbeamertemplate{itemize subitem}{\footnotesize\raisebox{0.3ex}{\tikz\fill[Gold] (0,0) circle (0.35ex);}}

% Suppress beamer's cosmetic underfull \hbox warnings on flushright/RaggedRight lines.
% Overfull warnings remain enforced.
\hbadness=10000

% Custom commands
\newcommand{\transitionslide}[2]{%
    {\setbeamercolor{normal text}{fg=SoftWhite,bg=DeepNavy}
    \begin{frame}[plain]\vfill\begin{center}
        {\huge\bfseries\textcolor{Gold}{#1}}\\[0.6cm]
        {\large\textcolor{LightGray}{#2}}
    \end{center}\vfill\end{frame}}
}

\title{\textbf{The Solow Paradox of Generative AI}}
\subtitle{Wages Flat, Postings Down, the Commons Hollowing Out}
\author{Audit synthesis}
\date{}

\begin{document}

% ---------- TITLE ----------
{\setbeamertemplate{footline}{}
\begin{frame}[plain]
\titlepage
\vspace{-0.6cm}
\begin{center}
\footnotesize\textcolor{WarmGray}{%
Humlum \& Vestergaard (2025), \emph{BFI WP 2025-56} \quad\textbar\quad
Liu, X.\ (2025), \emph{Working Paper}\\
Burtch, Lee \& Chen (2024), \emph{Scientific Reports} 14:10413 \quad\textbar\quad
Lyu et al.\ (2025), \emph{WWW Companion '25} (Acemoglu et al.)}
\end{center}
\end{frame}}

% ---------- OPENING PROVOCATION ----------
\begin{frame}[plain]
\vfill
\begin{center}
{\large\textcolor{WarmGray}{``You can see the computer age everywhere}}\\
{\large\textcolor{WarmGray}{but in the productivity statistics.''}}\\[0.2cm]
{\footnotesize\textcolor{WarmGray}{\textemdash{} Solow (1987), invoked by Humlum (2025) at p.\ 22.}}\\[1.0cm]
{\Huge\bfseries\textcolor{DeepRed}{$0\%$ \,\textbar\, $-19\%$ \,\textbar\, $-12\%$ \,\textbar\, $-29{,}013$}}\\[0.6cm]
{\large\textcolor{DeepNavy}{wage effect on incumbents \,\textbar\, entry postings (Y3) \,\textbar\, Stack Overflow traffic \,\textbar\, Wikipedia views/article}}\\[0.6cm]
{\Large\textcolor{WarmGray}{The aggregates show nothing. The pipeline and the commons say otherwise.}}
\end{center}
\vfill
\end{frame}

% ---------- FRAMEWORK ----------
\begin{frame}{Three places to look for displacement \textemdash{} only one shows zero}
\begin{center}
\renewcommand{\arraystretch}{1.05}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{3.6cm}>{\RaggedRight\arraybackslash}p{4.4cm}>{\RaggedRight\arraybackslash}p{4.2cm}@{}}
\toprule
\textbf{Where to look} & \textbf{What it measures} & \textbf{Signal} \\
\midrule
\textbf{Incumbent wages \& hours} & Pass-through of productivity to pay & \textcolor{Teal}{\textbf{Zero}} (Humlum 2025) \\
\textbf{New-hire demand} & Postings, hires, vacancy mix & \alert{$-12$\% to $-19$\%} (Liu 2025) \\
\textbf{Knowledge commons} & Q\&A traffic, edits, contributions & \alert{Hollowing} (Burtch 24; Lyu 25) \\
\bottomrule
\end{tabular}
\end{center}
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Stocks (incumbents) look fine. Flows (entry, contribution) do not.}
\end{flushright}
\end{frame}

% ---------- TRANSITION 1 ----------
\transitionslide{Incumbents look untouched}{Adoption is wide. Pass-through is narrow.}

% ============================================================
%  SECTION A \textemdash{} HUMLUM (DENMARK, PRECISE ZEROS)
% ============================================================

% ---------- HUMLUM: ADOPTION VS PASS-THROUGH ----------
\begin{frame}{Adoption ran from 47\% to 83\%. Pass-through to earnings was 3\textendash 7\%.}
\begin{center}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{5.4cm}p{3.0cm}p{3.0cm}@{}}
\toprule
\textbf{Margin} & \textbf{No encourage.} & \textbf{Encouraged} \\
\midrule
``Ever used'' AI chatbot & 47\% & \alert{83\%} \\
Daily use & 8\% & 21\% \\
Time savings (\% of work hours) & $\sim$2.0\% & 2.8\textendash 6.8\% \\
\textbf{Pass-through to earnings} & \textcolor{Teal}{\textbf{$\sim$3\%}} & \textcolor{Teal}{\textbf{$\sim$7\%}} \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.1cm}
Workers \emph{used} the tool. Time \emph{was} saved. Almost none of it landed in pay.\\
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Humlum \& Vestergaard (2025), Tables~1\textendash 2; Figs.\ 2, 4, 10. \(N\!\approx\!25{,}000\) workers, 11 occupations.}
\end{flushright}
\end{frame}

% ---------- HUMLUM: PRECISE ZEROS ----------
\begin{frame}{Across 11 exposed occupations, the dynamic DiD rules out wage effects above 1\%}
\begin{center}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{5.0cm}p{3.6cm}p{3.0cm}@{}}
\toprule
\textbf{Specification} & \textbf{CI on \% wage effect} & \textbf{F or N} \\
\midrule
Dynamic DiD (Fig.\ 7) & rules out $>$2\% & N $\approx$ 25K \\
Pooled DiD ``All'' (Fig.\ 8a) & rules out $>$1\% & 11 occs \\
Per-occupation pooled & rules out $>$6\% & each \\
Coworker IV (Tab.\ E.1) & confirms OLS & F\,=\,3{,}645 \\
Workplace-level DiD (Fig.\ 11) & rules out $>$1\% & wage bill \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.1cm}
Same answer in five different cuts. Adopters, encouraged adopters, daily users, early adopters, and high-adoption workplaces are all statistically indistinguishable from controls.\\
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Humlum \& Vestergaard (2025), Figs.\ 7\textendash 11; admin data through June 2024.}
\end{flushright}
\end{frame}

% ---------- TRANSITION 2 ----------
\transitionslide{But the door is closing on new hires}{What incumbents kept, entrants will not get.}

% ============================================================
%  SECTION B \textemdash{} LIU 2025 (LIGHTCAST POSTINGS)
% ============================================================

% ---------- LIU: DYNAMIC EFFECTS ----------
\begin{frame}{Job postings in AI-substitutable occupations fell 12\% on average \textemdash{} and the effect is widening}
\begin{center}
\renewcommand{\arraystretch}{1.0}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{4.4cm}p{3.0cm}p{3.6cm}@{}}
\toprule
\textbf{Horizon (post-ChatGPT)} & \textbf{$\Delta$ postings} & \textbf{SE / sig.} \\
\midrule
Pooled average & \alert{$-12.2\%$} & 0.036 (***) \\
Year 1 (2023) & $-6.4\%$ & $\sim$0.036 (*) \\
Year 2 (2024) & $-13.5\%$ & $\sim$0.036 (**) \\
\textbf{Year 3 (2025)} & \alert{\textbf{$-18.2\%$}} & 0.036 (***) \\
\bottomrule
\end{tabular}
\end{center}
Decile\,$\times$\,quarter FEs absorb exposure level; identified margin is \emph{substitutability}. Pre-trend Wald: F\,=\,0.83, p\,=\,0.66.
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Liu (2025), Table~1, Col.~3; 285M Lightcast postings, $N\!=\!6{,}753{,}930$ cells, 2018Q1\textendash 2025Q2.}
\end{flushright}
\end{frame}

% ---------- LIU: HETEROGENEITY ----------
\begin{frame}{Entry-level and administrative postings absorb the entire decline}
\begin{center}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{5.4cm}p{3.0cm}p{3.0cm}@{}}
\toprule
\textbf{Slice} & \textbf{$\Delta$ postings} & \textbf{Note} \\
\midrule
\textbf{Entry-level} & \alert{$-19\%$} & largest age effect \\
Mid-level & smaller & attenuates \\
Senior & near zero & insulated \\
\midrule
Administrative support & \alert{$-40\%$} & sectoral peak \\
Professional services & $-30\%$ & second largest \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.1cm}
The earnings of incumbent senior workers are stable. The \emph{path into} those jobs is not.\\
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Liu (2025), heterogeneity analysis; experience and 2-digit industry interactions.}
\end{flushright}
\end{frame}

% ---------- TRANSITION 3 ----------
\transitionslide{And the knowledge commons hollows out}{The training data degrades when its contributors leave.}

% ============================================================
%  SECTION C \textemdash{} BURTCH (STACK OVERFLOW), LYU (WIKIPEDIA)
% ============================================================

% ---------- BURTCH: TRAFFIC + REDDIT CONTROL ----------
\begin{frame}{Stack Overflow lost $\sim$1M visitors per day; Reddit's developer subs lost zero}
\begin{center}
\renewcommand{\arraystretch}{1.0}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{5.0cm}p{3.6cm}p{3.0cm}@{}}
\toprule
\textbf{Outcome} & \textbf{Stack Overflow} & \textbf{Reddit dev subs} \\
\midrule
Daily traffic (SCUL, Fig.\ 1) & \alert{$-12\%$} ($\sim$1M/day) & \textcolor{Teal}{\textbf{null}} \\
Question volume (DiD, Fig.\ 2) & \alert{$-10$ to $-15$/wk} & \textcolor{Teal}{\textbf{null}} \\
Hardest-hit topics (SO) & Python, JS, ReactJS, Pandas & \textcolor{Teal}{n/a} \\
\bottomrule
\end{tabular}
\end{center}
\textbf{Social fabric moderates.} LLMs substitute for pure information; communities with bonds (Reddit) absorb the shock, communities without them (Stack Overflow) bleed.
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Burtch, Lee \& Chen (2024), Figs.\ 1\textendash 4; SimilarWeb daily, Sept 2022\textendash Mar 2023.}
\end{flushright}
\end{frame}

% ---------- BURTCH: NEWER USERS EXIT, COMPLEXITY RISES ----------
\begin{frame}{Newer users leave first; the questions that remain are harder}
\begin{center}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{5.6cm}p{4.4cm}@{}}
\toprule
\textbf{Compositional shift, week 15 post-ChatGPT} & \textbf{Magnitude} \\
\midrule
Avg.\ posting-user account tenure & \alert{$+100$ to $+200$ days} \\
Avg.\ long-word frequency in questions & \alert{$+0.3$ to $+0.5$ units} \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.1cm}
\textbf{The juniors who left} were learning by asking. They lose mentorship and peer feedback \textemdash{} and they are precisely the cohort least able to verify a hallucinated LLM answer.\\
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Burtch, Lee \& Chen (2024), Figs.\ 5\textendash 6; replicated via Athey et al.\ (2021) matrix completion in Fig.\ S2.}
\end{flushright}
\end{frame}

% ---------- LYU: WIKIPEDIA ----------
\begin{frame}{Wikipedia articles substitutable by GPT lose 29{,}013 views \textemdash{} and the edits drop too}
\begin{center}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{4.6cm}p{3.6cm}p{3.0cm}@{}}
\toprule
\textbf{Outcome} & \textbf{DiD coefficient} & \textbf{Sig.} \\
\midrule
Pageviews / article-month & \alert{$-29{,}013$} (SE 6{,}904) & p $<$ 0.01 \\
Edits / article-month & \alert{$-6.88$} (SE 3.72) & p $<$ 0.10 \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.1cm}
Treatment = above-median cosine similarity to GPT-3.5 output. \(N\!=\!2{,}206\) articles, 24-month event window. The view drop is robust; the edit drop is suggestive but only marginally so.\\[0.1cm]
\textbf{The same loop as Stack Overflow.} Fewer readers $\Rightarrow$ fewer editors $\Rightarrow$ weaker training corpus.\\
\begin{flushright}
\footnotesize\textcolor{WarmGray}{Lyu et al.\ (2025), \emph{WWW Companion '25}, Table~1; Acemoglu among co-authors.}
\end{flushright}
\end{frame}

% ---------- DEVIL'S ADVOCATE ----------
\begin{frame}{Devil's advocate: aren't these all just the J-curve before the J turns up?}
\begin{block}{The strongest objection}
Brynjolfsson, Rock \& Syverson (2021): firms in the trough of a productivity J-curve sink complementary investments now that pay off later. Humlum (2025, p.\ 21) flags this; his most common new task is ``AI integration'' itself.
\end{block}
\textbf{Reply.} J-curve explains the zero on incumbent wages \textemdash{} not:
\begin{itemize}
\item Liu (2025): entry postings declined three years running ($-6.4, -13.5, -18.2\%$). A J-curve trough does not deepen monotonically.
\item Burtch (2024): contributor exit is \emph{compositional}; newer users leave. That selection cannot be reversed by waiting.
\item Lyu (2025): the commons that trains the models is meanwhile shrinking.
\end{itemize}
J-curve is incumbent-side optimism. The cost falls on entrants and on commons.
\end{frame}

% ---------- SYNTHESIS ----------
\begin{frame}{Stocks are stable. Flows are bleeding. The signal hides in the latter.}
\begin{center}
\renewcommand{\arraystretch}{1.05}
\begin{tabular}{@{}>{\columncolor{LightGray}}p{3.0cm}>{\RaggedRight\arraybackslash}p{3.4cm}>{\RaggedRight\arraybackslash}p{5.0cm}@{}}
\toprule
\textbf{Variable type} & \textbf{Signal} & \textbf{Empirical anchor} \\
\midrule
Incumbent stock & \textcolor{Teal}{precise zero} & Humlum: CI $<$ 1\% on wages \\
Entry flow & \alert{$-19\%$ Y3 entry} & Liu: $-6.4 \to -18.2\%$ over 3 yrs \\
Knowledge inflow & \alert{$-12\%$ traffic} & Burtch: SCUL on Stack Overflow \\
Knowledge upkeep & \alert{$-29$K views} & Lyu: DiD on Wikipedia \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.1cm}
The economic measurements that would call this an ``effect'' \textemdash{} earnings, hours, employment of incumbents \textemdash{} are exactly the variables that miss it. Look at flows.
\end{frame}

% ---------- CLOSING ----------
{\setbeamercolor{normal text}{fg=SoftWhite,bg=DeepNavy}
\setbeamertemplate{footline}{}
\begin{frame}[plain]
\vfill
\begin{center}
{\Huge\bfseries\textcolor{Gold}{The aggregates are clean.}}\\[0.6cm]
{\Large\textcolor{LightGray}{The pipeline and the commons are not.}}
\end{center}
\vfill
\end{frame}}

\end{document}
